---
title: "Synthetic overidentified full-sample DML-GMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic overidentified full-sample DML-GMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This example uses the canonical heterogeneous-cell data accompanying the
paper: 3,521 observations in all 1,400 row-column cells, with cell sizes from
one to four. There are two instruments and one target parameter, so the model
is overidentified. Four nuisance regressions are tuned and then refitted on
the complete sample. The tuning folds below choose hyperparameters; they are
not DML cross-fitting.

The full example is installed at
`system.file("examples", "synthetic_overidentified.R", package = "fullsampleDML")`.
The chunks are not evaluated while building the vignette because the four
tuning searches are intentionally more expensive than routine package checks.

```{r, eval=FALSE}
library(fullsampleDML)
library(mlr3)
library(mlr3tuning)
library(paradox)

data_path <- system.file("extdata", "synthetic_gmm.csv", package = "fullsampleDML")
data <- read.csv(data_path)
x_cols <- paste0("x", 1:5)
cluster_cols <- c("row_id", "col_id")
cv_mode <- "oneway" # change to "multiway" for strict tuning
```

The one-way mode automatically uses `col_id`, the dimension with 35 rather
than 40 clusters. Strict multiway mode constructs 3-by-3 Cartesian folds; no
training and validation observation shares either a row or a column cluster.

```{r, eval=FALSE}
make_learner <- function() {
  auto_tuner(
    tuner = tnr("grid_search", resolution = 10),
    learner = lrn("regr.glmnet", alpha = 0),
    resampling = rsmp("cv", folds = 5),
    measure = msr("regr.mse"),
    term_evals = 10,
    search_space = ps(
      lambda = p_dbl(log(1e-3), log(1e3), trafo = exp)
    )
  )
}

features <- function(frame) {
  stats::model.matrix(
    ~ (x1 + x2 + x3 + x4 + x5)^2 +
      I(x1^2) + I(x2^2) + I(x3^2) + I(x4^2) + I(x5^2) - 1,
    data = frame
  )
}

groups <- if (cv_mode == "multiway") {
  multiway_group_kfold(cluster_cols, 3, shuffle = TRUE, random_state = 2026)
} else {
  NULL
}

specs <- list(
  ell = nuisance_spec(make_learner(), "y", features, groups),
  r = nuisance_spec(make_learner(), "d", features, groups),
  m1 = nuisance_spec(make_learner(), "z1", features, groups),
  m2 = nuisance_spec(make_learner(), "z2", features, groups)
)
```

The two-component orthogonal score is

\[
\psi_i(\theta)=
\begin{pmatrix}\widetilde Z_{1i}\\\widetilde Z_{2i}\end{pmatrix}
(\widetilde Y_i-\theta\widetilde D_i).
\]

```{r, eval=FALSE}
score <- function(data, theta, nuisance) {
  predictions <- nuisance$predictions
  y_residual <- data$y - predictions$ell
  d_residual <- data$d - predictions$r
  z_residual <- cbind(
    data$z1 - predictions$m1,
    data$z2 - predictions$m2
  )
  z_residual * (y_residual - theta[1] * d_residual)
}

jacobian <- function(data, theta, nuisance) {
  predictions <- nuisance$predictions
  d_residual <- data$d - predictions$r
  z_residual <- cbind(
    data$z1 - predictions$m1,
    data$z2 - predictions$m2
  )
  matrix(colMeans(z_residual * d_residual), ncol = 1)
}

result <- fit_gmm(
  data, score, theta_start = 0, cluster_cols = cluster_cols,
  nuisance_fit = mlr3_nuisance_fit(specs), jacobian = jacobian,
  weight_type = "PSD", covariance_type = "PSD"
)

summary(result)
result$n_observations # 3521
result$n_cells        # 1400
result$n_eff          # 35
result$nuisance$group_columns$ell
```

Software compatibility alone does not establish statistical validity. Users
remain responsible for the rate, complexity, orthogonality, and dependence
conditions in the accompanying paper.
